New and simplified manual controls for projection and slice tours, with application to exploring classification boundaries in high dimensions

### 2022-09-05 paper/paper.Rmd
software section
Reworking the slicing section

To better explore the space we combine the manual controls for the projection with manual controls for slicing. A slice is defined by a projection, a center point that is anchoring it in the high-dimensional space and the slice thickness $h$ [@slicetour].

@@

In the case of a single orthogonal direction on the projection plane we can pick a sequence of center points in steps along that direction to move the slice and fully cover the data space. This no longer works in higher-dimensional spaces, and we can think of picking one direction and shifting the slice along the component orthogonal to the projection plane. In practice we define the center point explicitly and calculate the slice display as explained in @slicetour. The current implementation allows the user to individually enter the components of the vector defining the center point, and updates the display by jumping to the new slice.

XXX Equation for arbitrary center for slice; and move the "the implementation ..." into the software section once that is constructed; and "potential adjustments for future" into the section on desirable controls

@@

In addition it is also useful to interactively change the slice thickness, in particular when selecting the preferred thickness $h$ (also called the slice radius) for exploring the input data. This was implemented as a slicer, and to allow the user to efficiently explore the interesting range we can provide the slider range as an input. For guidance the estimates of the number of points inside the slice as a function of the original sample size $N$ and the number of dimensions $p$ from @sectionpursuit can be used: in case of a uniform distribution inside a sphere of radius $R$ a slice with thickness $h$ will contain $N_S$ points, with 
\begin{equation}
N_S = \frac{N}{2} \left(\frac{h}{R}\right)^{p-2} \left(p - (p-2)\left(\frac{h}{R}\right)^{2}\right).
\label{eq:count}
\end{equation}

@@

@@

Explain `radial_tour`

>>>>

To better explore the space we combine the manual controls for the projection with manual controls for slicing. A slice is a section of the data that is defined by a projection, a center point that is anchoring it in the high-dimensional space and the slice thickness $h$ [@slicetour]. A data point is inside the slice if its orthogonal distance from the projection plane (passing through the center point) is below the thickness $h$. This orthogonal distance is computed in terms of the component that is normal on the projection plane. For $\mathbf{x}_i$ a $p$ dimensional data point and $\mathbf{c}$ the center point (in the same $p$ dimensional space) we compute the orthogonal distance as
\begin{equation}
v_i^2 = ||\mathbf{x}_i' - \mathbf{c}'||^2 = \mathbf{x}_i'^2 + \mathbf{c}'^2 - 2 \mathbf{x}_i'\cdot\mathbf{c}' ,
\label{eq:slice}
\end{equation}
with $\mathbf{c}' = \mathbf{c} - (\mathbf{c}\cdot \mathbf{a}_1) \mathbf{a}_1 - (\mathbf{c}\cdot \mathbf{a}_2 )\mathbf{a}_2$, $\mathbf{x}_i' = \mathbf{x}_i - (\mathbf{x}_i\cdot \mathbf{a}_1) \mathbf{a}_1 - (\mathbf{x}_i\cdot \mathbf{a}_2) \mathbf{a}_2$ and $\mathbf{a}_k, k=1,2 (=d)$ denoting the columns of the projection matrix, $\mathbf{A}=(\mathbf{a}_1, \mathbf{a}_2)$.

@@

In the case of a single orthogonal direction on the projection plane we can pick a sequence of center points $\mathbf{c}$ in steps along that direction to move the slice and fully cover the data space. This no longer works in higher-dimensional spaces, and we can think of picking one direction and shifting the slice along the component orthogonal to the projection plane. In practice we define the center point $\mathbf{c}$ explicitly in terms of all $p$ components and calculate the slice display following Eq. \ref{eq:slice}.

@@

In addition it is also useful to interactively change the slice thickness $h$ (also called the slice radius), in particular to find the preferred value for exploring the input data. This was implemented as a slider, and to allow the user to efficiently explore the interesting range we provide the slider range as an input. For guidance the estimates of the number of points inside the slice as a function of the original sample size $N$ and the number of dimensions $p$ from @sectionpursuit can be used: in case of a uniform distribution inside a sphere of radius $R$ a slice with thickness $h$ will contain $N_S$ points, with 
\begin{equation}
N_S = \frac{N}{2} \left(\frac{h}{R}\right)^{p-2} \left(p - (p-2)\left(\frac{h}{R}\right)^{2}\right).
\label{eq:count}
\end{equation}

@@

XXX copied from the slicing part to be weaved in below:

The current implementation allows the user to individually enter the components of the vector defining the center point, and updates the display by jumping to the new slice.

@@

The @tourr package provides numerous types of tours. An additional function has been constructed that will generate a sequence of projections that will decrease the coordinates for $V_m$ to $\boldsymbol{0}$ and back to the original values. It is applicable for any projection dimension, $d$.

### 2022-09-05 paper/paper.tex
Reworking the slicing section

To better explore the space we combine the manual controls for the
projection with manual controls for slicing. A slice is defined by a
projection, a center point that is anchoring it in the high-dimensional
space and the slice thickness \(h\) \citep{slicetour}.

@@

In the case of a single orthogonal direction on the projection plane we
can pick a sequence of center points in steps along that direction to
move the slice and fully cover the data space. This no longer works in
higher-dimensional spaces, and we can think of picking one direction and
shifting the slice along the component orthogonal to the projection
plane. In practice we define the center point explicitly and calculate
the slice display as explained in \citet{slicetour}. The current
implementation allows the user to individually enter the components of
the vector defining the center point, and updates the display by jumping
to the new slice.

XXX Equation for arbitrary center for slice; and move the ``the
implementation \ldots{}'' into the software section once that is
constructed; and ``potential adjustments for future'' into the section
on desirable controls

@@

In addition it is also useful to interactively change the slice
thickness, in particular when selecting the preferred thickness \(h\)
(also called the slice radius) for exploring the input data. This was
implemented as a slicer, and to allow the user to efficiently explore
the interesting range we can provide the slider range as an input. For
guidance the estimates of the number of points inside the slice as a
function of the original sample size \(N\) and the number of dimensions
\(p\) from \citet{sectionpursuit} can be used: in case of a uniform
distribution inside a sphere of radius \(R\) a slice with thickness
\(h\) will contain \(N_S\) points, with \begin{equation}
N_S = \frac{N}{2} \left(\frac{h}{R}\right)^{p-2} \left(p - (p-2)\left(\frac{h}{R}\right)^{2}\right).
\label{eq:count}
\end{equation}

>>>>

To better explore the space we combine the manual controls for the
projection with manual controls for slicing. A slice is a section of the
data that is defined by a projection, a center point that is anchoring
it in the high-dimensional space and the slice thickness \(h\)
\citep{slicetour}. A data point is inside the slice if its orthogonal
distance from the projection plane (passing through the center point) is
below the thickness \(h\). This orthogonal distance is computed in terms
of the component that is normal on the projection plane. For
\(\mathbf{x}_i\) a \(p\) dimensional data point and \(\mathbf{c}\) the
center point (in the same \(p\) dimensional space) we compute the
orthogonal distance as \begin{equation}
v_i^2 = ||\mathbf{x}_i' - \mathbf{c}'||^2 = \mathbf{x}_i'^2 + \mathbf{c}'^2 - 2 \mathbf{x}_i'\cdot\mathbf{c}' ,
\label{eq:slice}
\end{equation} with
\(\mathbf{c}' = \mathbf{c} - (\mathbf{c}\cdot \mathbf{a}_1) \mathbf{a}_1 - (\mathbf{c}\cdot \mathbf{a}_2 )\mathbf{a}_2\),
\(\mathbf{x}_i' = \mathbf{x}_i - (\mathbf{x}_i\cdot \mathbf{a}_1) \mathbf{a}_1 - (\mathbf{x}_i\cdot \mathbf{a}_2) \mathbf{a}_2\)
and \(\mathbf{a}_k, k=1,2 (=d)\) denoting the columns of the projection
matrix, \(\mathbf{A}=(\mathbf{a}_1, \mathbf{a}_2)\).

@@

In the case of a single orthogonal direction on the projection plane we
can pick a sequence of center points \(\mathbf{c}\) in steps along that
direction to move the slice and fully cover the data space. This no
longer works in higher-dimensional spaces, and we can think of picking
one direction and shifting the slice along the component orthogonal to
the projection plane. In practice we define the center point
\(\mathbf{c}\) explicitly in terms of all \(p\) components and calculate
the slice display following Eq. \ref{eq:slice}.

@@

In addition it is also useful to interactively change the slice
thickness \(h\) (also called the slice radius), in particular to find
the preferred value for exploring the input data. This was implemented
as a slider, and to allow the user to efficiently explore the
interesting range we provide the slider range as an input. For guidance
the estimates of the number of points inside the slice as a function of
the original sample size \(N\) and the number of dimensions \(p\) from
\citet{sectionpursuit} can be used: in case of a uniform distribution
inside a sphere of radius \(R\) a slice with thickness \(h\) will
contain \(N_S\) points, with \begin{equation}
N_S = \frac{N}{2} \left(\frac{h}{R}\right)^{p-2} \left(p - (p-2)\left(\frac{h}{R}\right)^{2}\right).
\label{eq:count}
\end{equation}
